---
title: "How to Create Inset Maps"
author: "Timothée Giraud"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Create Inset Maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```




**`mf_inset*()`** functions create map insets.
`mf_inset_on()` starts the inset and `mf_inset_off()` closes it. 


If map layers have to be plotted after the inset (i.e after `mf_inset_off()`), please use `add = TRUE`.  
It is not possible to plot an inset within an inset.  
It is possible to plot anything (base plots) within the inset, not only map layers.  



## Size and Position

There is 2 solution for defining the size and position of an inset.   

The first solution is based on `x`, `pos` and `cex` arguments:  

- `x` is an sf object, the width/height ratio of the inset will match the width/height ratio of `x` bounding box;  
- `pos` must be one of "bottomleft", "left", "topleft", "top", "bottom", "bottomright", "right" or"topright";   
- `cex` defines the share of the map width covered by the inset (0.5 means 50%).   

In the following example we create an inset for a specific municipality in the topright corner of the map. 

```{r, fig.height=6, fig.width=5, results = FALSE}
library(mapsf)
# set a theme
mf_theme("default", mar = c(0, 0, 0, 0))
# data import
mtq <- mf_get_mtq()
# Selection of a target municipality to display in the inset
mtq_target <- mtq[c(26), ]
# Display the base map
mf_map(mtq)
# display the target on the main map
mf_map(mtq_target, add = TRUE, col = "tomato")
# open the inset
mf_inset_on(x = mtq_target, pos = "topright", cex = .4)
# display the target municipality
mf_map(mtq_target, add = FALSE, col = "tomato")
# close the inset
mf_inset_off()
```

Once an inset is opened, map elements can be added to it.  
Here we add a title, a shadow, a scale bar, other Martinique municipalities and a frame.  

```{r, fig.height=7.58, fig.width=6.25, results = FALSE}
# set a theme
mf_theme("dark")
# Selection of a target municipality to display in the inset
mtq_target <- mtq[c(26), ]
# Display the base map
mf_map(mtq)
# display the target on the main map
mf_map(mtq_target, add = TRUE, col = "tomato")
# open the inset
mf_inset_on(x = mtq_target, pos = "topright", cex = .4)
# center on on the target munucipality
mf_init(mtq_target)
# display all municipalities
mf_map(mtq, add = TRUE)
# display the target municipality shadow
mf_shadow(mtq_target, add = TRUE)
# display the target municipality
mf_map(mtq_target, add = TRUE, col = "tomato")
# display a title
mf_title("Saint-Anne",
  pos = "left", tab = TRUE,
  cex = .9, line = 1, inner = TRUE
)
# display a scale bar
mf_scale(size = 2)
# dispaly a frame around the inset
box()
# close the inset
mf_inset_off()
# display the map layout of the main map
mf_title("Martinique Municipalities")
# display a scale bar
mf_scale(size = 5)
# display credits
mf_credits(txt = "T. Giraud, 2021")
```

The second solution to set the size and position of an inset is to use the `fig` argument.   
`fig` coordinates (xmin, xmax, ymin, ymax) are expressed as fractions of the mapping space (i.e. excluding margins).  
`mf_inset_on(fig = c(0, 0.25, 0, 0.25))` will create an inset on the bottomleft corner of the map and will cover a quarter of the map in width and height.   

In the following example we create an inset for a specific municipality in the bottom left corner of the map. 

```{r, fig.height=6, fig.width=5, results = FALSE}
# set a theme
mf_theme("default", mar = c(0, 0, 0, 0))
# Display the base map
mf_map(mtq)
# open the inset
mf_inset_on(fig = c(0, 0.25, 0, 0.25))
# display a municipality
mf_map(mtq[9, ])
# add a frame
box()
# close the inset
mf_inset_off()
```


## World Map Inset

It is easy to create an inset with a world map (see `mf_worldmap()`) using `x = "worldmap"`. 

```{r, fig.height=6, fig.width=5, results = FALSE}
mf_map(mtq)
mf_inset_on(x = "worldmap")
mf_worldmap(mtq)
mf_inset_off()
```



## Non-cartographic Insets
Insets can be used for regular plots. In the next map we create an inset with an histogram of the median income distribution using the map classification. 

```{r, fig.height=7.58, fig.width=6.25, results = FALSE}
# create classes
bks <- mf_get_breaks(x = mtq$MED, nbreaks = 5, breaks = "quantile")
bks <- round(bks, digits = -2)
# define a color palette
pal <- hcl.colors(n = 5, palette = "Dark Mint", rev = TRUE)
# set a theme
mf_theme("candy")
fg <- mf_theme()$fg
# Create the map
mf_map(
  x = mtq, var = "MED", type = "choro",
  pal = pal, breaks = bks, leg_pos = NA
)
# start the inset
mf_inset_on(fig = c(0.75, .95, 0.84, .99))
# draw the histogram
par(mar = c(0, 0, 1.7, 0))
hist(mtq$MED,
  breaks = bks, col = pal, border = fg, axes = FALSE, labels = "",
  xlab = "", ylab = "", main = ""
)
axis(
  side = 1, at = bks, las = 2, tick = FALSE, line = -.9,
  cex.axis = .7, col.axis = fg
)
title("Median Income\nin euros",
  cex.main = .8, col.main = fg,
  font.main = 1, adj = 0
)
# close the inset
mf_inset_off()
# Add map layout
mf_title("Wealth in Martinique, 2015", pos = "left")
mf_scale(5)
mf_credits(paste0(
  "Sources: Insee and IGN, 2018\n",
  "mapsf ", packageVersion("mapsf")
))
```

