---
title: "How to Use Themes"
author: "Timothée Giraud"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Use Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 4,
  fig.height = 4
)
```


**`mf_theme()`** sets a map theme. A theme is a set of graphical parameters that are applied to maps created with `mapsf`. These parameters are the figure margins, background and foreground colors and `mf_title()` options. `mapsf` offers some builtin themes. It's possible to modify an existing theme or to start a theme from scratch. The map theme can also be set with `mf_init()` and `mf_export()`.

## Builtin themes

Here are the builtin themes. We have added a grey background color to make the margins more visible. 
```{r, echo = FALSE}
par(bg = "grey")
suppressPackageStartupMessages(library(mapsf))
themes <- c(
  "default", "ink", "dark", "agolalight", "candy",
  "darkula", "iceberg", "green", "nevermind", "jsk", "barcelona"
)
mtq <- mf_get_mtq()
for (i in themes) {
  mf_theme(i)
  mf_map(mtq)
  mf_title(paste0('"', i, '"'))
  box(which = "figure", col = "white")
}
```

The following table displays the parameters of these themes.

```{r, echo = FALSE}
a <- NULL
for (i in themes) {
  ax <- mf_theme(i)
  a <- rbind(a, paste0(ax))
}
a <- as.data.frame(a)
names(a) <- c("name", "bg", "fg", "mar", "tab", "pos", "inner", "line", "cex", "font")
# pipe tables by default
knitr::kable(a)
```

## How to modify an existing theme

It is possible to modify an existing theme. In this example we use the "default" theme and modify some title parameters. 
```{r, echo = FALSE}
par(bg = "grey")
```
```{r}
library(mapsf)
mtq <- mf_get_mtq()
mf_theme("default", tab = FALSE, font = 4)
mf_map(mtq)
mf_title('Mofified "default" theme')
```

## How to create a new theme

It is possible to create a new theme. 

```{r}
mf_theme(
  bg = "lightblue", fg = "tomato1", mar = c(0, 2, 0, 2),
  tab = TRUE, inner = TRUE, line = 1.3, pos = "center",
  cex = 1.1, font = 3
)
mf_map(mtq)
mf_title("New theme")
```



