---
title: "How to Use Themes"
author: "Timothée Giraud"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Use Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 4,
  fig.height = 4, 
  fig.path = "fig/wo_theme_"
)
```


**`mf_theme()`** sets a map theme. A theme is a set of graphical parameters that are applied to maps created with `mapsf`.   
These parameters are:

- figure margins, 
- background, foreground and highlight colors, 
- default sequential and qualitative palettes,
- title options (position, size, banner...).

`mapsf` offers some builtin themes. It's possible to modify an existing theme or to start a theme from scratch.
Themes are persistent across maps produced by `mapsf` (e.g. they survive a `dev.off()` call). 

## Builtin themes

Here are the builtin themes.   

```{r, echo = FALSE, fig.width=8, fig.height=8}
suppressPackageStartupMessages(library(mapsf))
themes <- list(
  'base' = 'base', 
  'sol_dark' = 'sol_dark',
  'sol_light' = 'sol_light',
  'grey' = 'grey',
  'mint' = 'mint',
  'dracula' = 'dracula'
)
opar <-par(mfrow=c(3,4))
mtq <- mf_get_mtq()
for (i in 1:length(themes)) {
  mf_theme(themes[[i]])
  mf_map(mtq)
  mf_map(mtq, "POP", "prop", inches = .15, leg_pos = 'topright')
  mf_layout(paste0("Theme: ", names(themes)[[i]]))
  mf_map(mtq, "MED", "choro", leg_frame = FALSE)
  mf_title("Sequential Palette")
}
```


## How to modify an existing theme

It is possible to modify an existing theme. In this example we use the "base" theme and modify some title parameters. 

```{r}
library(mapsf)
mtq <- mf_get_mtq()
mf_theme("base", title_banner = FALSE, title_font = 4)
mf_map(mtq)
mf_title('Mofified "default" theme')
```

## How to create a new theme

It is possible to create a new theme from scratch. 

```{r}
mf_theme(
    mar          = c(0.5, 0.5, 2.25, 0.5)+-.4,
    title_tab    = FALSE,
    title_pos    = "left",
    title_inner  = FALSE,
    title_line   = 1.75,
    title_cex    = 1.25,
    title_font   = 4,
    title_banner = FALSE,
    foreground   = "#F1F0E9",
    background   = "white",
    highlight    = "#E9762B",
    pal_quali    = "Dark 3",
    pal_seq      = colorRampPalette(c("#F1F0E9", "#41644A", "#0D4715"))
  )
mf_map(mtq, "MED", "choro")
mf_title("New theme")
```


It is also possible to assign a theme to a variable. 


```{r}
blue_theme <- mf_theme("base", background = "lightblue")
green_theme <- mf_theme("base", background = "lightgreen")
mf_theme(blue_theme)
mf_map(mtq)
mf_title("Blue Theme")
mf_theme(green_theme)
mf_map(mtq)
mf_title("Green Theme")
```
